---
title: "Project 2 Code Base Submission Dataset 1"
author: "Long Lin"
format: html
editor: visual
---

## Overview

For this project, I used three different datasets from the Week 5 Discussion 5A post. With these three datasets, I prepared each of them by creating a .csv file and importing the data. Then I worked on tidying the data, and performing an analysis on the dataset. I also made sure that the code within the Quarto Markdown file is reproducible in a clean environment. I used a similar process to what we did in Assignment 5A with the Airline Delays, as I feel like that is very similar assignment to this.

## Dataset 1: 

Birth Rate by Countries posted by Brandon Chanderban

source: https://raw.githubusercontent.com/longflin/DATA-607/refs/heads/main/Project%202/Birth_Rates_of_Countries.csv

```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
library(gt)

birth_rate_url <- "https://raw.githubusercontent.com/longflin/DATA-607/refs/heads/main/Project%202/Birth_Rates_of_Countries.csv"

birth_rate_df <- read_csv(
  file = birth_rate_url,
  show_col_types = FALSE,
  progress = FALSE,
  skip = 4
)

head(birth_rate_df)
```
I converted the data from a wide format to a long format using pivot_longer in the following code chunk.
```{r}
long_birth_rate_df <- birth_rate_df |>
  pivot_longer(
    cols = c("1960":"2022"),
    names_to = "Year",
    values_to = "Birth Rate"
  )
head(long_birth_rate_df, 70)
```

I replaced missing values with 0 with the following code chunk.

```{r}
long_birth_rate_remove_na_df <- long_birth_rate_df |>
  mutate(`Birth Rate` = replace_na(`Birth Rate`, 0))

head(long_birth_rate_remove_na_df, 70)
```
I filtered the data for US only data using the following code chunk.  Then I displayed the US only data in a table.
```{r}
US_only_df <- long_birth_rate_remove_na_df |>
  filter(`Country Name` == 'United States')
head(US_only_df, 70)

US_only_df |>
  gt() |>
  cols_hide(columns = c(`Country Name`, `Country Code`, `Indicator Name`, `Indicator Code`)) |>
  tab_header(
    title = "Birth rates in the US (per 1,000 people)",
  )
```
I created a line chart for the US birth rate data using the following code chunk.  This was done to make it easier to see trends.
```{r}
library(ggplot2)

ggplot(US_only_df, aes(x = as.numeric(Year), y = `Birth Rate`)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(1960, 2022, by = 10)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
    title = "US Birth Rate Trends Over Time",
    subtitle = "Annual births per 1,000 persons",
    x = "Year",
    y = "Birth Rate",
    color = "Region"
  )
```
From this plot, we are able to see that the US has been on a downward trend in crude birth rates between the year 1960 and 2022.
