---
title: "Project 2 Code Base Submission Dataset 3"
author: "Long Lin"
format: html
editor: visual
---

## Overview

For this project, I used three different datasets from the Week 5 Discussion 5A post. With these three datasets, I prepared each of them by creating a .csv file and importing the data. Then I worked on tidying the data, and performing an analysis on the dataset. I also made sure that the code within the Quarto Markdown file is reproducible in a clean environment. I used a similar process to what we did in Assignment 5A with the Airline Delays, as I feel like that is very similar assignment to this.

## Dataset 3: 

World GDP by Country: 1960-2022 posted by Sinem Kilicdere

source: https://raw.githubusercontent.com/longflin/DATA-607/refs/heads/main/Project%202/World%20GDP%20by%20Country%201960-2022.csv

```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
library(gt)

world_GDP_url <- "https://raw.githubusercontent.com/longflin/DATA-607/refs/heads/main/Project%202/World%20GDP%20by%20Country%201960-2022.csv"

world_GDP_df <- read_csv(
  file = world_GDP_url,
  show_col_types = FALSE,
  progress = FALSE
)

head(world_GDP_df)
```
I fixed the missing values in this dataset by replacing them with 0 using the following code chunk.
```{r}
world_GDP_df <- world_GDP_df |>
  mutate(across("1960":"2022", ~replace_na(.x, 0)))

head(world_GDP_df)
```
I converted the data from a wide data format to a long data format using pivot_longer.
```{r}
long_world_GDP_df <- world_GDP_df |>
  pivot_longer(
    cols = c("1960":"2022"),
    names_to = "Year",
    values_to = "GDP"
  )
head(long_world_GDP_df, 70)
```
Next I filtered the data for US only data using the following code chunk.  I also converted the GDP value to billions by dividing the value by 1000000000.
```{r}
US_only_GDP_df <- long_world_GDP_df |>
  filter(`Country` == 'United States')
head(US_only_GDP_df, 70)

US_only_GDP_df <- US_only_GDP_df |>
  mutate(GDP = as.numeric(GDP) / 1000000000)

US_only_GDP_df |>
  gt() |>
  cols_hide(columns = c(`Country`, `Country Code`)) |>
  tab_header(
    title = "US GDP by Year (billions)",
  )
```
I created a line chart for the US GDP data using the following code chunk.  This was done so that it is easier to see trends.
```{r}
library(ggplot2)

ggplot(US_only_GDP_df, aes(x = as.numeric(Year), y = GDP)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(1960, 2022, by = 10)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
    title = "US GDP Over Time",
    subtitle = "GDP (billions)",
    x = "Year",
    y = "GDP (millions)",
    color = "Region"
  )
```

The plot for the US' GDP trends upwards between 1960 and 2022.

Next I created a data frame of China's GDP data by using a filter.  I also converted the GDP into billions for China's values by dividing the GDP value by 1000000000.  Then I created a table of the values for the Year and GDP columns.
```{r}
China_only_GDP_df <- long_world_GDP_df |>
  filter(`Country` == 'China')
head(US_only_GDP_df, 70)

China_only_GDP_df <- China_only_GDP_df |>
  mutate(GDP = as.numeric(GDP) / 1000000000)

China_only_GDP_df |>
  gt() |>
  cols_hide(columns = c(`Country`, `Country Code`)) |>
  tab_header(
    title = "China GDP by Year (billions)",
  )
```
I created a line chart for China's GDP data using the following code chunk.  This was done so that it is easier to see trends.
```{r}
library(ggplot2)

ggplot(China_only_GDP_df, aes(x = as.numeric(Year), y = GDP)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(1960, 2022, by = 10)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
    title = "China GDP Over Time (billions)",
    x = "Year",
    y = "GDP (billions)",
    color = "Region"
  )
```

The plot for China's GDP also trends upwards between 1960 and 2022.

Finally, I created a plot for both US and China's GDP data in order to make it easier to compare both countries together.
```{r}
US_only_GDP_df <- US_only_GDP_df |>
  mutate(Source = "US")

China_only_GDP_df <- China_only_GDP_df |>
  mutate(Source = "China")

combined_df <- bind_rows(US_only_GDP_df, China_only_GDP_df) 

ggplot(combined_df, aes(x = as.numeric(Year), y = GDP, color = Source, group = Source)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(1960, 2022, by = 10)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
    title = "China GDP Over Time (billions)",
    x = "Year",
    y = "GDP (billions)",
    color = "Region"
  )
```
Combining the plots for the US and China shows that the US' GDP grows at a faster rate than China's GDP.