---
title: "Assignment 5B ELO Calculations"
author: "Long Lin"
format: html
editor: visual
---

## Overview

For this assignment, I will use the chess tournament information from Project 1 and calculate each player's expected score and the difference from their actual score. From there, I will list the five players who overperformed the most relative to their expected score, as well as the five players that underperformed the most relative to their expected score. I think the formula given in the reference video is a good start and should be reasonable for this assignment.

source: https://raw.githubusercontent.com/longflin/DATA-607/refs/heads/main/Assignment%205B/tournamentinfo.txt

## The Process

I will start with creating a data frame from the chess tournament information provided in project 1.
```{r}
library(tidyverse)
library(gt)

# 1. Load the raw text
raw_lines <- readLines("https://raw.githubusercontent.com/longflin/DATA-607/refs/heads/main/Assignment%205B/tournamentinfo.txt")

# 2. Remove dashes and empty lines
clean_lines <- raw_lines[!grepl("^-+$", raw_lines) & nzchar(raw_lines)]

# 3. Skip the header (first 2 lines of the remaining text)
data_lines <- clean_lines[3:length(clean_lines)]

# 4. Separate the two-row structure
# Every odd line is a "Name" row, every even line is a "ID/Rating" row
name_rows <- data_lines[seq(1, length(data_lines), by = 2)]
info_rows <- data_lines[seq(2, length(data_lines), by = 2)]

# 5. Parse the Name Rows (Fixed Width)
# Column positions: Pair Num (1-5), Name (9-40), Total Pts (36-40), Rounds (42-end)
names_df <- data.frame(
  pair_num = as.numeric(str_sub(name_rows, 1, 5)),
  player_name = str_trim(str_sub(name_rows, 9, 40)),
  total_pts = as.numeric(str_sub(name_rows, 42, 44)),
  stringsAsFactors = FALSE
)

# 6. Parse the Info Rows (Extracting State and Pre-Rating)
info_df <- data.frame(
  state = str_trim(str_sub(info_rows, 1, 5)),
  uscf_id = str_extract(info_rows, "\\d{8}"),
  pre_rating = as.numeric(str_extract(info_rows, "(?<=R: )\\s*\\d+")),
  stringsAsFactors = FALSE
)

# 7. Merge names with info
tournament_df <- cbind(names_df, info_df)

head(tournament_df, 100)
```
For the round data, I created a data frame called opponents_df and stored the rounds as well as the opponent_pre_rating for each round.
```{r}
# Extract the string containing the rounds
rounds_raw <- str_sub(name_rows, 42, -1)

# Extract just the digits (Opponent Pair Numbers)
# We use str_extract_all to get every number in the row
opponents_list <- str_extract_all(rounds_raw, "(?<=[WLD])\\s*\\d+")

# Convert the list into a data frame where each row is a game
opponents_df <- data.frame(
  pair_num = rep(tournament_df$pair_num, lengths(opponents_list)),
  opponent_id = as.numeric(unlist(opponents_list))
)

#Grab the opponent's pre rating from the tournament_df
opponents_df$opponent_pre_rating <- tournament_df$pre_rating[match(opponents_df$opponent_id, tournament_df$pair_num)]

head(opponents_df)
```
Once I parsed the data from the tournamentinfo.txt file, I created a matches data frame by taking the opponents_df and adding the pre_rating value associated with each pair_num in the opponents_df using a left_join.
```{r}
matches_df <- opponents_df |>
  left_join(
    tournament_df |> select(pair_num, pre_rating),
    by = "pair_num"
  ) |>
  relocate(pre_rating, .after = pair_num)
head(matches_df, 10)
```
In order to calculate the expected score, I'll use the formula from singingbanana's video "The ELO Rating System for Chess and Beyond".

singingbanana. (2019, February 15). The Elo Rating System for Chess and Beyond [Video]. YouTube. https://www.youtube.com/watch?v=AsYfbmp0To0

P(A wins) = 1/(1+(10^((RB-RA)/400)))

I used the formula to add another column of the expected_score for each match.
```{r}
matches_df <- matches_df |>
  mutate(expected_score = 1 / (1 + (10^((opponent_pre_rating - pre_rating) / 400))))

head(matches_df)
```
Next I calculated the expected points for each player by adding up all the expected scores for each pair_num.
```{r}
expected_points_df <- matches_df |>
  group_by(pair_num) |>
  summarise(expected_points = sum(expected_score), .groups = "drop")

head(expected_points_df)
```
Next I calculated the differential between the player's total points and their expected points.
```{r}
differential_df <- tournament_df |>
  left_join(expected_points_df, by = "pair_num") |>
  mutate(
    differential = total_pts - expected_points
  )

head(differential_df)
```
In order to get the top 5 overperformers, I used dplyr's arrange function to arrange the list in descending order and then used the head function to grab the top 5.
```{r}
overperformed_df <- differential_df |>
  arrange(desc(differential))

top5_overperformers <- head(overperformed_df, 5)

top5_overperformers |>
  gt() |>
  cols_label(
    pair_num = "Pair Number",
    player_name = "Player Name",
    total_pts = "Total Pts",
    state = "State",
    uscf_id = "USCF ID",
    pre_rating = "Pre-rating",
    expected_points = "Expected Pts",
    differential = "Differential"
  ) |>
  tab_header(
    title = "Top 5 Overperformers",
  )
```
In order to get the top 5 underperformers, I used dplyr's arrange function to arrange the list in the default ascending order and then used the head function to grab the top 5.
```{r}
underperformed_df <- differential_df |>
  arrange(differential)

top5_underperformers <- head(underperformed_df, 5)

top5_underperformers |>
  gt() |>
  cols_label(
    pair_num = "Pair Number",
    player_name = "Player Name",
    total_pts = "Total Pts",
    state = "State",
    uscf_id = "USCF ID",
    pre_rating = "Pre-rating",
    expected_points = "Expected Pts",
    differential = "Differential"
  ) |>
  tab_header(
    title = "Top 5 Underperformers",
  )
```

Finally  I checked my work by hand calculating to see if it was correct.  I used Gary Hua for the calculation.

```{r}
tournament_df |>
  filter(pair_num == 1)

opponents_df |>
  filter(pair_num ==1)

#opponent 39
expected_score39 = 1 / (1 + (10^((1436 - 1794) / 400)))

#opponent 21
expected_score21 = 1 / (1 + (10^((1563 - 1794) / 400)))

#opponent 18
expected_score18 = 1 / (1 + (10^((1600 - 1794) / 400)))

#opponent 14
expected_score14 = 1 / (1 + (10^((1610 - 1794) / 400)))

#opponent 7
expected_score7 = 1 / (1 + (10^((1649 - 1794) / 400)))

#opponent 12
expected_score12 = 1 / (1 + (10^((1663 - 1794) / 400)))

#opponent 4
expected_score4 = 1 / (1 + (10^((1716 - 1794) / 400)))

expected_gary_hua = sum(expected_score39, expected_score21, expected_score18, expected_score14, expected_score7, expected_score12, expected_score4)

expected_gary_hua

expected_points_df |> filter(pair_num == 1)
```
Hand calculating for Gary Hua resulted in an expected score of 5.161574 while the algorithm resulted in the same number so it is correct.




## Conclusion

The ELO formula provided by singingbanana's youtube video was reliable in calculating each player's expected score.  In the end I was able to find the top 5 overperformers and the top 5 underperformers.  I was also able to verify the accuracy of the expected score by comparing GARY HUA's score with my own hand calculations.
